#CASA0005 Week 4 Gender Inequality Homework - Hasan Guler

###load libraries that we need
```{r}

library(sf)
library(tmap)
library(here)
library(sp)
library(dplyr)
library(janitor)
library(tidyverse)
library(countrycode)
```
###load the global gender inequality data
```{r}

hdr_data <- read.csv(here::here("week4_hm_data", "HDR21-22_Composite_indices_complete_time_series.csv"))
```

###load our spatial data, which is the boundries of countries in the world
```{r}
world_data <- st_read(here::here("week4_hm_data", "World_Countries__Generalized_.shp"))
```

###global gender data has 1008 variables so choose only those required

Selecting only required fields, which are gender inequality index in 2019 and gender inequality index in 2010

```{r}
gender_data_columns <- hdr_data %>%
  dplyr::select(c("iso3", "country", "region", "gii_2010","gii_2019"))
  
```
###rearrange our subset gender inequality data to make it ready for the join process

1. Change the name of iso3 column to standardize our data making the same name with world data
2. Standardize all the column names
3. Creating a new column called difference where gii_2019 - gii_2010
4. Create a new column by means of converting iso3 into iso2


```{r}

gender_data <- gender_data_columns %>%
  dplyr::rename(iso = "iso3") %>%
  clean_names() %>%
  mutate(diff_2019_2010 = gii_2019-gii_2010) %>%
  mutate(iso_code=countrycode(iso, origin ='iso3c', destination = 'iso2c'))

 


```
###rearrange the column names of world data since all column names are in uppercase

```{r}
world_data_sub <- world_data %>%
  clean_names()
class(world_data_sub)
```
###Join our data to have a spatial world data with gii from UNDP data

```{r}

joined_data <- world_data_sub %>%
  left_join(., 
            gender_data,
            by = c("iso" = "iso_code")) %>%
  drop_na(diff_2019_2010)

  
```

```{r}
join_fr<-joined_data %>%
  filter(iso=="FR")
```

#displaying on a map (just for testing purpose)

The map does not represent the data effectively since it is just created to test the joined data.
```{r}

tm_shape(joined_data) + 
tm_polygons("diff_2019_2010", 
        title="Gender Inequality Difference",
        alpha = 0.5)
```








